---
title: "01-exercises"
author: "Ryan Guzalowski"
date: "April 6, 2016"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  


### Solution 

```{r}
f <- function(x) {
  
  return(sapply(1:length(x), function(i) {prod(x[-i])}))
}

x <- c(1,5,2,8)
f(x)
```

## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers except for: 
   multiples of 3 replaced by "Fizz"? 
   multiples of 5 replaced by "Buzz" 
   multiples of 3 and 5 replaced by "FizzBuzz""
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz


### Solution

```{r}
f <- function(x) {
  
  retVal <- x
  
  multiplesOf3 <- (x%%3 == 0)
  retVal[multiplesOf3] <- "Fizz"
  
  multiplesOf5 <- (x%%5 == 0)
  retVal[multiplesOf5] <- "Buzz"
  
  multiplesOf3and5 <- (x%%3 == 0 & x%%5 == 0)
  retVal[multiplesOf3and5] <- "FizzBuzz"
  
  return(retVal)
}

x <- 1:20
f(x)
```
