---
title: "NYCFlights: Arrival Delay Logictic Model"
author: "Ryan Guzalowski"
date: "5/2/17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(readr)
library(magrittr)
library(lubridate)
library(stringr)
library(data.table)

flights <- read_csv("data/flights.csv")
planes   <- read_csv("data/planes.csv")
airports <- read_csv("data/airports.csv") 
weather  <- read_csv("data/weather.csv")

flightsDT  <- fread("data/flights.csv") 
planesDT   <- fread("data/planes.csv")  %>% setkey(tailnum)
airportsDT <- fread("data/airports.csv")  %>% setkey(faa) 
weatherDT  <- fread("data/weather.csv") %>%   setkey(origin,year,month,day,hour)

YX <- flightsDT 
YX %<>% merge( planesDT, all.x = TRUE, by='tailnum', suffixes=c('','.pl') )
YX %<>% merge( weatherDT, all.x = TRUE, by=c('origin','year','month','day','hour'), suffixes=c('','.we') )
YX %<>% merge( airportsDT, all.x = TRUE, by.x='origin', by.y='faa', suffixes=c('','.orig') )
YX %<>% merge( airportsDT, all.x = TRUE, by.x='dest', by.y='faa', suffixes=c('','.dest') )

```


## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`

**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }

lf <- function(x) {
  return (1/(1+exp(-x)))
}

p1 <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
p1 + stat_function(fun = lf) + xlim(-10,10) + annotate("text", x = -5, y = .75, label = "1/(1+exp(-x))", size = 10)

lfinv <- function(y) {
  return (-log((1-y)/y))
}

p2 <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
p2 + stat_function(fun = lfinv) + xlim(0,1) + annotate("text", x = .2, y = 2.5, label = "ln(x/(1-x))", size = 10)

# Testing that they are invereses. Input == output
test <- c(.1, .2, .4, .7, .9)
test %>% lf() %>% lfinv() %>% print()

```


# PART A:

Using the rectangular data that you created from the earlier assignment and following the example from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.   

```{r, run}

# reducing sample size, choosing predictors
yx <- YX %>% sample_n(10000)
y <- "arr_delay"
xs <- c('month', 'dep_delay', 'carrier', 'air_time'
        ,'distance', 'year.pl', 'type', 'engine'
        ,'wind_dir', 'wind_speed', 'wind_gust', 'precip'
        ,'pressure', 'visib', 'lat', 'lon'
        ,'lat.dest', 'lon.dest')
yx <- yx[, c(y,xs), with = FALSE]

# setting up arrival delay to be factor of 0 (<22min) or 1 (>=22 min)
# r uses second factor to try and fit, so we make (>=22 min) == 1
yx$arr_delay[yx$arr_delay < 22] <- 0
yx$arr_delay[yx$arr_delay >= 22] <- 1
yx$arr_delay <- factor(yx$arr_delay)
levels(yx$arr_delay)

# training simple model
fit1 <- glm(arr_delay ~ dep_delay, family=binomial, data=yx)
summary(fit1)

plot(yx$dep_delay, as.numeric(yx$arr_delay)-1,
     main = "Only using dep_delay as Predictor")
curve(predict(fit1,data.frame(dep_delay=x),type="resp"),add=TRUE)


# training model with all predictors included
fit2 <- glm(arr_delay ~ ., family=binomial, data=yx)
summary(fit2)


```


# Question A:

Is this a good model?

AIC went from 4300 in the model with only the dep_delay (simple model) to 2520 using
all the predictors, so it is an improvement.

# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

```{r}

# make the training set. choosing predictors that we would know only in advance
yx <- YX %>% sample_n(10000)
y <- "arr_delay"
xs <- c("origin", "month", "day", "hour", "sched_dep_time", "sched_arr_time",
        "carrier", "air_time", "distance", "type", "engines", "seats",
        "lat.dest", "lon.dest", "alt.dest")
yxb <- yx[, c(y,xs), with = FALSE]

# Set up arrival delay specifications
yxb$arr_delay[yxb$arr_delay < 22] <- 0
yxb$arr_delay[yxb$arr_delay >= 22] <- 1
yxb$arr_delay <- factor(yxb$arr_delay)
levels(yxb$arr_delay)

# training simple model
fitb1 <- glm(arr_delay ~ air_time, data = yxb, family = "binomial")
summary(fitb1)

plot(yxb$air_time, as.numeric(yxb$arr_delay)-1,
     main = "Using only air_time as predictor")
curve(predict(fitb1, data.frame(air_time=x),type="resp"),add=TRUE)

# training model with all predictors
fitb2 <- glm(arr_delay ~ ., data = yxb, family = "binomial")
summary(fitb2)



```


# Question B:

Is this a good model?

AIC is now 7400, much worse than when we had the day-of information. It is a improvement over the simple model of using only air_time though, which gave an AIC of 9600. As shown by the graph though, even the most statistically significant predictor fails to obviously show a trend