---
title: "Classification Metrics"
author: "Ryan Guzalowski"
date: "May 9, 2017"
output: html_document
---

**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions"}

library(dplyr)

TPR <- function(y, yhat) {
  truePos <- yhat[y == TRUE & yhat == TRUE] %>% length()
  obsPos <- y[y == TRUE] %>% length()
  return(truePos/obsPos)
}

FPR <- function(y, yhat) {
  falsePos <- yhat[yhat == TRUE & y == FALSE] %>% length()
  obsNeg <- y[y == FALSE] %>% length()
  return(falsePos/obsNeg)
}

TNR <- function(y, yhat) {
  trueNeg <- yhat[y == FALSE & yhat == FALSE] %>% length()
  obsNeg <- y[y == FALSE] %>% length()
  return(trueNeg/obsNeg)
}

FNR <- function(y, yhat) {
  falseNeg <- yhat[yhat == FALSE & y == TRUE] %>% length()
  obsPos <- y[y == TRUE] %>% length()
  return(falseNeg/obsPos)
}

sensitivity <- function(y, yhat) {
  return(TPR(y, yhat))
}

specificity <- function(y, yhat) {
  return(TNR(y, yhat))
}

recall <- function(y, yhat) {
  return(TPR(y, yhat))
}

precision <- function(y, yhat) {
  truePos <- yhat[y == TRUE & yhat == TRUE] %>% length()
  predictedPos <- yhat[y==TRUE] %>% length()
  return(truePos/predictedPos)
}

prevalence <- function(y, yhat) {
  actualPos <- y[y == TRUE] %>% length()
  total <- length(y)
  return(actualPos/total)
}

accuracy <- function(y, yhat) {
  truePos <- yhat[y == TRUE & yhat == TRUE] %>% length()
  trueNeg <- yhat[y == FALSE & yhat == FALSE] %>% length()
  total <- length(y)
  return((truePos+trueNeg)/total)
}

kappa <- function(y, yhat) {
  
  po <- accuracy(y, yhat)

  n <- length(y)
  TP <- yhat[y == TRUE & yhat == TRUE] %>% length()
  FN <- yhat[yhat == FALSE & y == TRUE] %>% length()
  TN <- yhat[y == FALSE & yhat == FALSE] %>% length()
  FP <- yhat[yhat == TRUE & y == FALSE] %>% length()
  
  pT <- (TP+FN)/n * (TP+FP)/n
  pF <- (FP+TN)/n * (FN+TN)/n
  
  pe <- pT+pF
  
  return((po-pe)/(1-pe))
}


funs <- list(TPR, FPR, TNR, FNR, sensitivity, specificity, recall, precision,
             prevalence, accuracy, kappa)

funNames <- list("True Positive Rate", "False Positive Rate",
                 "True Negative Rate", "False Negative Rate",
                 "Sensitivity", "Specificity", "Recall", "Precision",
                 "Prevalence", "Accuracy", "Kappa")

```


## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}

logitModel <- readRDS( "logitModel.rds")
y <- logitModel$y
yhat <- logitModel$yhat

for (i in 1:length(funs)) {
  print(paste(funNames[[i]], ":", funs[[i]](y,yhat) %>% round(digits=2)))
}

```