---
title: "Assingment NYC Flight Data"
author: "Ryan Guzalowski"
date: "April 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 

Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
library(readr)
library(dplyr)

# reading flights data from its csv file using package readr
flights <- read_csv("data/flights.csv", col_types = cols())
glimpse(flights)
```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}

#histogram for arrival delays
hist(flights$arr_delay,
     breaks = 100,
     main = "Distribution of Flight Arrival Delay Times",
     xlab = "Delay Time [min]",
     xlim = c(-100, 300))

#histogram for departure delays
hist(flights$dep_delay,
     breaks = 100,
     main = "Distribution of Flight Departure Delay Times",
     xlab = "Delay Time [min]",
     xlim = c(-100, 300))

```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}

#isolating data
dates <- as.Date(flights$time_hour)
arrivalDelay <- flights$arr_delay

#scatter plot
plot(arrivalDelay, dates,
     main = "Flight Delays in 2013",
     xlab = "Delay Time [min]",
     ylab = "Date",
     pch = 20,
     cex = .05)
```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}

# first group the data
byOriginCarrier <- group_by(flights, origin, carrier)

#then use summarise() to isolate grouped cols
numFlts <- summarise(byOriginCarrier, count = n())
glimpse(numFlts)

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}

#importing the other data tables
airports <- read_csv("data/airports.csv", col_types = cols())
weather <- read_csv("data/weather.csv", col_types = cols())
planes <- read_csv("data/planes.csv", col_types = cols())

#joining the tables together
masterTable <- flights %>% select(year:time_hour)
masterTable <- masterTable %>% left_join(planes, by = "tailnum")
masterTable <- masterTable %>% left_join(airports, by = c("origin" = "faa"))
masterTable <- masterTable %>% left_join(airports, by = c("dest" = "faa"),
                                         suffix = c(".origin", ".dest"))
masterTable <- masterTable %>% left_join(weather, by = c("month", "day", "origin", "hour"))
glimpse(masterTable)

```



