---
title: "Resampling"
author: "Ryan Guzalowski"
date: "May 9, 2017"
output: html_document
---

**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file**

## Resampling

`x` is a random variable. We want to not only know what the `mean(x)` is but want to calculate the uncertainty of `mean(x)`.  Measuring the uncertainty requires repeated measurements of `mean(x)`.

- Calculate the mean of `x`.
- Calculte the `sd( mean(x) )` using the **using 10-fold resampling**.  Create your own folds, show your work.
- Calculte the `sd( mean(x) )` using the **using 10 bootstrap sample**.  Create your own folds, show your work. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(dplyr)
```

### 10 FOLD:
```{r "sd_cv"}
set.seed(1) 
x <- runif(20,1,20)

# 10 Fold resampleing fun
sd_cv <- function(x) {
  
  n <- length(x)
  foldSize <- n/10
  
  start <- 1
  end <- foldSize
  means <- rep(0, 10) #init
  for (i in 1:10) {
    means[i] <- x[-(start:end)] %>% mean() #takes out fold and gets mean of sample
    start <- start + foldSize #primes indicies for next round
    end <- end + foldSize
  }
  
  return (sd(means)) #returns sd of all means collected
}

print(paste("10 FOLD ::: Mean:", x %>% mean() %>% round(digits=2), ",",
            "SD:", sd_cv(x) %>% round(digits=2)))

```


### 10 BOOTSTRAP:
```{r "sd_boot"}

sd_boot <- function(x) {
  
  n <- length(x)
  
  means <- rep(0, 10) #init
  for (i in 1:10) {
    bootStrap <- base::sample(x, n, replace=TRUE) #samples with replacement
    means[i] <- bootStrap %>% mean()
  }
  
  return (sd(means)) #returns sd of all means collected
}


print(paste("10 BOOTSTRAP ::: Mean:", x %>% mean() %>% round(digits=2), ",",
            "SD:", sd_boot(x) %>% round(digits=2)))
```



